---
title: "Fitting synthetic data using epidp"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fitting_synthetic_data_using_epidp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(epidp)
library(ggplot2)
library(dplyr)
library(magrittr)
library(purrr)
library(tidyr)
options(mc.cores=4)
```


In this vignette, we show how `epidp` can be used to generate then fit to synthetically
generated infection data. Throughout, we assume a generating process of the form:

\begin{equation}
I_t \sim \text{Poisson}(R_t \Lambda_t),
\end{equation}

where $\Lambda:=\sum_{s=1}^t \omega_s I_{t-s}$.


## Step function in $R_t$
We first generate case data assuming a step function for $R_t$.
```{r}
rt_fun = function(t){
  if(t <= 60)
    R = 2
  else if (t <= 90)
    R = 0.5
  else
    R = 1
  R
}

# simulation parameters
nt <- 200
mean_si <- 6.5
sd_si <- 4.03
i_0 <- 10

# data frame of outputs
epidemic_df <- simulate_renewal_epidemic(rt_fun, nt, mean_si, sd_si, i_0)

# plot
transform_factor <- 300
epidemic_df %>%
  select(-c(w_dist, lambda_t)) %>%
  mutate(R_t = R_t * transform_factor) %>% 
  pivot_longer(c(i_t, R_t)) %>%
  ggplot(aes(x = t, y = value, colour = name)) +
  geom_line() +
  scale_y_continuous(
    name = "Incidence (i_t)",
    sec.axis = sec_axis(~ . / transform_factor, name = "Reproduction Number (R_t)")
  ) +
  labs(x = "Time") +
  theme_minimal() +
  scale_color_brewer("Series", palette = "Dark2")
```

We now use a Stan version of EpiFilter to estimate the maximum a posteriori estimates
of $R_t$ and overlay these on top of the actual values. Note, these estimates do
not have uncertainty associated with them but the benefit of this is that estimation
is instantaneous. The estimates are close to the actual $R_t$ values after an initial
period when case counts are low.
```{r}
# fit model
fit <- fit_epifilter(
  N=length(epidemic_df$i_t),
  C=epidemic_df$i_t,
  w=epidemic_df$w_dist,
  is_sampling=FALSE,
  as_vector=FALSE
)

# plot
R <- fit$par$R
epidemic_df %>% 
  mutate(estimated=R) %>% 
  rename(true=R_t) %>% 
  select(t, estimated, true) %>% 
  pivot_longer(c(estimated, true)) %>% 
  ggplot(aes(x=t, y=value)) +
  geom_line(aes(colour=name)) +
  scale_color_brewer("R_t", palette = "Dark2") +
  ylab("R_t")
```


## Sinusoidal function in $R_t$
We now generate case data assuming a sinusoidal $R_t$.
```{r}
# define sinusoidal Rt
rt_fun <- function(t) {
  1.3 + 1.2 * sin(4 * (pi / 180) * t)
}
nt <- 200
mean_si <- 6.5
sd_si <- 4.03
i_0 <- 10

# data frame of outputs
epidemic_df <- simulate_renewal_epidemic(rt_fun, nt, mean_si, sd_si, i_0)

# plot
transform_factor <- 150
epidemic_df %>%
  select(-c(w_dist, lambda_t)) %>%
  mutate(R_t = R_t * transform_factor) %>% 
  pivot_longer(c(i_t, R_t)) %>%
  ggplot(aes(x = t, y = value, colour = name)) +
  geom_line() +
  scale_y_continuous(
    name = "Incidence (i_t)",
    sec.axis = sec_axis(~ . / transform_factor, name = "Reproduction Number (R_t)")
  ) +
  labs(x = "Time") +
  theme_minimal() +
  scale_color_brewer("Series", palette = "Dark2")
```

We now use a Stan version of EpiFilter to estimate the $R_t$ profile.
```{r}
# fit model
fit <- fit_epifilter(
  N=length(epidemic_df$i_t),
  C=epidemic_df$i_t,
  w=epidemic_df$w_dist,
  iter=200
)

# look at MCMC summaries
print(fit, c("sigma", "R"))
```

We now overlay the estimated $R_t$ versus the actual values. The estimated $R_t$ values
coincide reasonably with the true values.
```{r}
# extract posterior quantiles
R_draws <- rstan::extract(fit, "R")[[1]]
lower <- apply(R_draws, 2, function(x) quantile(x, 0.025))
middle <- apply(R_draws, 2, function(x) quantile(x, 0.5))
upper <- apply(R_draws, 2, function(x) quantile(x, 0.975))

# plot
epidemic_df %>% 
  mutate(
    lower=lower,
    middle=middle,
    upper=upper
  ) %>% 
  select(t, R_t, lower, middle, upper) %>% 
  ggplot(aes(x=t, y=R_t)) +
  geom_line(colour="red") +
  geom_ribbon(aes(ymin=lower, ymax=upper), fill="blue", alpha=0.4) +
  geom_line(aes(y=middle))
```


