---
title: "Fitting real COVID-19 case data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(epidp)
library(ggplot2)
library(dplyr)
library(magrittr)
library(purrr)
library(tidyr)
```

# Estimating $R_t$ for Bogota using only the case data
We first load the COVID-19 data for Colombia and graph it.
```{r}
# load data
data("covid_colombia_cases_deaths_mobility")

# plot
covid_colombia_cases_deaths_mobility %>%
  pivot_longer(c(cases, deaths)) %>%
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  facet_grid(vars(name), vars(city),
    scales = "free"
  ) +
  xlab("Date") +
  ylab("Count")
```
We first estimate $R_t$ for Bogota using only the case data using optimisation to give us a quick set of estimates.
```{r}
df_bogota <- covid_colombia_cases_deaths_mobility %>%
  filter(city == "Bogota")

# generate serial interval for COVID-19 based on reasonable mean, sd
mean_si <- 6.5
sd_si <- 4.03
w <- generate_vector_serial(nrow(df_bogota), mean_si, sd_si)

# fit using optimisation
fit <- fit_epifilter(
  N = nrow(df_bogota),
  C = df_bogota$cases,
  w = w,
  is_sampling = FALSE,
  as_vector = FALSE
)

# plot
R <- fit$par$R
df_bogota %>%
  mutate(Rt = R) %>%
  select(date, Rt, cases) %>%
  filter(date >= as.Date("2020-04-01")) %>%
  pivot_longer(-date) %>%
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  scale_color_brewer("R_t", palette = "Dark2") +
  ylab("R_t") +
  facet_grid(vars(name), scales = "free")
```
We now fit using a fully Bayesian framework which outputs uncertainty.
```{r}
fit <- fit_epifilter(
  N = nrow(df_bogota),
  C = df_bogota$cases,
  w = w,
  is_sampling = TRUE,
  iter = 50,
  chains = 1 # as CRAN does not allow multiple cores
)

# extract posterior quantiles
R_draws <- rstan::extract(fit, "R")[[1]]
lower <- apply(R_draws, 2, function(x) quantile(x, 0.025))
middle <- apply(R_draws, 2, function(x) quantile(x, 0.5))
upper <- apply(R_draws, 2, function(x) quantile(x, 0.975))

# plot
df_bogota %>%
  mutate(
    lower = lower,
    middle = middle,
    upper = upper
  ) %>%
  select(date, lower, middle, upper) %>%
  filter(date >= as.Date("2020-04-01")) %>%
  ggplot(aes(x = date)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "blue", alpha = 0.6) +
  geom_line(aes(y = middle), colour = "blue") +
  geom_hline(yintercept = 1, linetype = 2, colour = "orange") +
  xlab("Date") +
  ylab("R_t")
```

# Probing the drivers of $R_t$ using mobility data
We assume a relationship between workplace mobility $m_t$ and $R_t$ of the form:

$$
\log(R_t) = \alpha + \beta m_t + \epsilon_t,
$$

where $\epsilon_t$ represents the components of $R_t$ unrelated to workplace mobility.

We now fit this model using `epidp`.

```{r}
X <- tibble(
  cons = rep(1, nrow(df_bogota)),
  m = df_bogota$workplaces
) %>%
  mutate(
    m = scale(m)[, 1]
  ) %>%
  as.matrix()

fit <- fit_epifilter_covariates(
  N = nrow(df_bogota),
  C = df_bogota$cases,
  w = w,
  X = X,
  is_sampling = TRUE,
  iter = 50, # probably too few iterations
  chains = 1 # as CRAN does not allow multiple cores; should run with more cores
)

print(fit, "beta[2]")
```
This negative association probably is a result of individuals responding to the COVID-19 pandemic conditions or governmental policy.

# Determining drivers of $R_t$ for Bogota using financial time series
```{r}
data("bogota_financial_time_series")
```

Smooth the weekly time series
```{r}
daily_dates <- seq(
  min(bogota_financial_time_series$date),
  max(bogota_financial_time_series$date), by = "day")

merchants <- unique(bogota_financial_time_series$merch_category)
for(i in seq_along(merchants)) {
  df_short <- bogota_financial_time_series %>% 
    filter(merch_category==merchants[i])
  spendamountusd_interpolated <- approx(
    df_short$date,
    df_short$spendamt, xout = daily_dates)$y
  countamountusd_interpolated <- approx(df_short$date, df_short$nb_transactions, xout = daily_dates)$y
  df_daily <- tibble(
      date=daily_dates,
      spendamountusd=spendamountusd_interpolated,
      countamountusd=countamountusd_interpolated
      ) %>% 
    mutate(merch_type=merchants[i])
  
  if(i == 1)
    big_df <- df_daily
  else
    big_df <- big_df %>% bind_rows(df_daily)
}

df_both <- big_df %>% 
  left_join(covid_colombia_cases_deaths_mobility) %>% 
  filter(city=="Bogota")
```

Here, we demonstrate how we can investigate whether the (smoothed) daily number of transactions in Bogota is associated with $R_$. We pick the "Grocery Stores/Supermarkets" category here because this is likely a particularly high contact shop.

```{r}
df_supermarkets <- df_both %>% 
  filter(merch_type=="Grocery Stores/Supermarkets")

X <- tibble(
  cons = rep(1, nrow(df_bogota)),
  m = df_supermarkets$countamountusd
) %>%
  mutate(
    m = scale(m)[, 1]
  ) %>%
  as.matrix()

options(mc.cores=4)
fit <- fit_epifilter_covariates(
  N = nrow(df_supermarkets),
  C = df_supermarkets$cases,
  w = w,
  X = X,
  is_sampling = TRUE,
  iter = 200,
  chains = 4 
)

print(fit, "beta[2]")
```
Plotting $R_t$ using these estimates.
```{r}
# extract posterior quantiles
R_draws <- rstan::extract(fit, "R")[[1]]
lower <- apply(R_draws, 2, function(x) quantile(x, 0.025))
middle <- apply(R_draws, 2, function(x) quantile(x, 0.5))
upper <- apply(R_draws, 2, function(x) quantile(x, 0.975))

# plot
df_supermarkets %>%
  mutate(
    lower = lower,
    middle = middle,
    upper = upper
  ) %>%
  select(date, lower, middle, upper) %>%
  filter(date >= as.Date("2020-04-01")) %>%
  ggplot(aes(x = date)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "blue", alpha = 0.6) +
  geom_line(aes(y = middle), colour = "blue") +
  geom_hline(yintercept = 1, linetype = 2, colour = "orange") +
  xlab("Date") +
  ylab("R_t")
```


