---
title: "Fitting synthetic data including covariates"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(epidp)
library(ggplot2)
library(dplyr)
library(magrittr)
library(purrr)
library(tidyr)
options(mc.cores=4)
```

In this document, we explore how the incorporation of covariate information affects
estimation of $R_t$.

## Added Gaussian noise around a sinusoidal covariate-driven mean $R_t$
```{r}
# define sinusoidal Rt with noise
rt_fun <- function(t, x) {
  x[1] * exp(x[2])
}
nt <- 200
t <- 1:nt
f <- 1.3 + 1.2 * sin(4 * (pi / 180) * t)
g <- vector(length = nt)
g[1] <- rnorm(1, 0, 1)
rho <- 0.8
for(i in 2:length(g)) {
  g[i] <- rho * g[i - 1] + rnorm(1, 0, 0.1)
}
X <- matrix(c(f, g), nrow = length(f), ncol = 2)

# simulation parameters
mean_si <- 6.5
sd_si <- 4.03
i_0 <- 10

# data frame of outputs
epidemic_df <- simulate_renewal_epidemic(rt_fun, nt, mean_si, sd_si, i_0, X)

# plot
epidemic_df %>%
  mutate(f=f) %>% 
  select(-c(w_dist, lambda_t)) %>%
  pivot_longer(c(R_t, f)) %>%
  ggplot(aes(x = t, y = value, colour = name)) +
  geom_line() +
  scale_y_continuous(
    name = "Incidence (i_t)",
    sec.axis = sec_axis(~ . / 1000, name = "Reproduction Number (R_t)")
  ) +
  labs(x = "Time", colour = "Variable") +
  theme_minimal()
```
We first try to estimate $R_t$ without covariate information.
We now use a Stan version of EpiFilter to estimate the $R_t$ profile.
```{r}
# fit model
fit <- fit_epifilter(
  N=length(epidemic_df$i_t),
  C=epidemic_df$i_t,
  w=epidemic_df$w_dist,
  iter=200
)

# look at MCMC summaries
print(fit, c("sigma", "R"))
```

We now overlay the estimated $R_t$ versus the actual values. The estimated $R_t$ values coincide reasonably with the true values.
```{r}
# extract posterior quantiles
R_draws <- rstan::extract(fit, "R")[[1]]
lower <- apply(R_draws, 2, function(x) quantile(x, 0.025))
middle <- apply(R_draws, 2, function(x) quantile(x, 0.5))
upper <- apply(R_draws, 2, function(x) quantile(x, 0.975))

# plot
epidemic_df %>% 
  mutate(
    lower=lower,
    middle=middle,
    upper=upper
  ) %>% 
  select(t, R_t, lower, middle, upper) %>% 
  ggplot(aes(x=t, y=R_t)) +
  geom_line(colour="red") +
  geom_ribbon(aes(ymin=lower, ymax=upper), fill="blue", alpha=0.4) +
  geom_line(aes(y=middle))
```

